### 2018-05-04 chapters/04-killings.tex
appdx 3; uamk models

Our analysis tests the sensitivity of 40 variables on a sample of 177 countries from 1945 to 2013. We find that GDP \textit{per capita}, democratic regimes, and the post-Cold War period are negatively associated with mass killings, whereas previous political turmoil increases the risk of such violence. In contrast, the distributed random forest algorithm indicates that years since the last episode of mass violence, GDP \textit{per capita}, urban population, openness to trade, and the number of military personnel make the greatest contribution to our models’ out-of-sample predictive power. While we discuss the importance of these findings in more detail later in this article, a few notes are imperative to make here. First, in some ways, these findings confirm previous research–unstable countries are more likely to witness the regime employ atrocity \citep[e.g.,]{goldsmith2013forecasting,harff2003no,krain1997state}.  However, many of the factors scholars often cite as observable indicators of such instability--regime transitions, coups d'état, the presence of militias, etc.--are not good proxies for instability. Thus, policymakers may be looking for the incorrect signs of impending atrocity when seeking to prevent its onset. Second, these findings raise concerns about policy options for preventing violence against civilians. If our conclusion is that unstable countries are violent, then preventing atrocity likely requires significant investments of time and resources in state-building, which is often politically and practically unfeasible \citep{doyle2006making}. Lastly, our findings are robust to a myriad of specifications, functional forms, and additional tests as we describe below. This analysis contributes significantly to the political violence literature by highlighting the parsimonious nature of the logic behind government atrocity and clearing away much of the empirical clutter surrounding this conclusion.

@@

We employ two methods to test the robustness of the potential determinants of state-led violence. First, we use a variant of extreme bounds analysis (EBA) to check the statistical significance of 35 explanatory variables cited in the academic literature.\footnote{The complete list of variables included in this article is available in the online appendix. Our selection procedure was as follows. We included all variables that appeared in at least two two quantitative papers on mass killings which employed logit or probit models in a global sample. We considered only published articles in our list.} Researchers have employed the EBA to assess the sensitivity of the determinants of civil war \citep{hegre2006sensitivity}, coups d'état \citep{gassebner2016expect}, democratisation \citep{gassebner2013extreme}, economic growth \citep{levine1992sensitivity, sala1997just, sturm2005determinants}, nuclear deterrence \citep{bell2015examining}, political repression \citep{hafner2005right}, and state corruption \citep{serra2006empirical}. The method is particularly useful when there is no consensus about which covariates belong in the ``true'' regression model \citep[178]{sala1997just} and scholars worry that omitted or unnecessary predictors could bias the parameter estimates \citep[60]{angrist2008mostly, clarke2005phantom, elwert2014endogenous, spector2011methodological}. By combining a large set of regression models into a single posterior distribution, the EBA reduces the uncertainty associated with the \textit{ad hoc} selection of predictors and provides an intuitive way of assessing the relative strength of the statistical findings \citep{leamer1985sensitivity, sala1997just}.

@@

In this paper, we follow his advice and consider the whole range of values of $CDF(0)$. We choose to use the whole distribution because the aggregate $CDF(0)$ allows us to move away from a binary indicator of robustness, the one we obtain with Leamer's extreme bounds, and present the estimations with their appropriate degrees of confidence \citep[179]{sala1997just}\footnote{As \citet[179]{sala1997just} describes, ``if 95 percent of the density function for the estimates of $\beta_{1}$ lies to the right of zero and only 52 percent of the density function for $\beta_{2}$  lies to the right of zero, one will probably think of variable 1 as being more likely to be correlated with [a dependent variable $Y$] than variable 2.''}. Our main focus is the percentage of the variable's cumulative distribution function that is smaller or greater than zero. We do not assume that the CDF has a normal distribution: Many of our coefficients do not follow a normal distribution, so the generic model provides a better fit to our data. We specify our models as follows:

@@

A different set of variables reach significance if we only consider mass killings that occur during civil wars. We ran the models on three indicators of violent armed conflict. The first is the civil war measure by the Uppsala Conflict Database Program  (\citeyear{allansson2017organized,gleditsch2002armed}). Here, two variables meet our criteria for robustness, Post-Cold War years and conflicts with territorial aims, both negatively correlated with mass killing onset. When we limit our analysis to cases marked as civil wars by the Correlates of War project, years since the last mass killing episode, previous riots, former human rights abuses, and ethnic diversity appear have a positive effect on the outcome. Rather surprisingly, the presence of militias has a negative effect on genocides and politicides. Finally, our model using ethnic civil war shows similar results: Militias have a negative impact on the likelihood of mass killings during ethnic conflict, as do conflicts over territorial aims. This finding is in contradiction with previous research \citep{koren2017means} and is deserving of further exploration. Overall, these results suggest two takeaways. First, the determinants of mass killing onset during peacetime and wartime are different from each other. While there are more restraints on atrocities during peacetime, these restraints fail once fighting begins. This raises particularly interesting questions about the pacifying nature of democracy. Second, how civil wars are classified significantly alters the results produced. This raises concerns for scholars about making claims based on only one data set.

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=\textwidth, height=9cm]{images/drfdpp4.pdf}
\caption{ Partial Dependence Plots -- Mass Killings during Civil Wars (Cederman et al. data)}
\label{fig:drfdpp4}
\end{center}
\end{figure}

@@

As a last set of robustness models, we estimate the same regressions using Harff's (\citeyear{harff2003no}) indicator of genocide and politicide. No variable appear as significant in our EBA models for genocide or politicide onset in peacetime. When we limit our sample to civil war years, the Post-Cold War period is again negatively correlated with the outcome when using the Correlates of War data set. Additionally, excluded population has a negative sign in more than 90\% of the models using both Correlates of War’s and Cederman et al's (\citeyear{cederman2010ethnic}) indicators of conflict. Displaced population also has a negative effect in the Correlates of War data set. During ethnic conflicts, our dummy variable for political assassinations has a negative impact on the onset of genocides. Overall, we conclude then that the significant covariate of genocide and politicide onset differ significantly from those of more general forms of government mass violence. Finally, the machine learning models are comparable to the ones we present above, with a similar set of variables appearing in the random forest estimations. The complete results are available in the appendix.

@@

\subsubsection{Complete Results}

Here we present a series of histograms with the distribution of coefficients of all variables of the main EBA model. There are 36 variables in total, of which seven are robust: Log GDP per capita, post-Cold War period, onset and ongoing civil wars (measured by the UCDP), previous riots, ethnic diversity and the squared term of the Polity IV index.

@@

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{Base variables} &  &  &  &  &  \\
Log GDP per capita & -0.0091 & 0.0052 & 76.055 & 0.9335 & 226707 \\
 &  &  &  &  &  \\
\textit{Additional variables} &  &  &  &  &  \\
Post-Cold War years & -0.0133 & 0.0085 & 72.845 & 0.9472 & 35614 \\
UCDP civil war onset & 0.0529 & 0.0321 & 52.378 & 0.9441 & 20854 \\
Previous riots & 0.0140 & 0.0100 & 56.242 & 0.9216 & 35614 \\
UCDP ongoing civil war & 0.0172 & 0.0115 & 65.652 & 0.9092 & 20854 \\
Ethnic diversity (ELF) & 0.0184 & 0.0137 & 56.674 & 0.9050 & 35614 \\
Polity IV squared & -0.0002 & 0.0001 & 61.206 & 0.9031 & 35614 \\ \hline
\end{tabular}
\caption{Extreme Bounds Analysis -- Mass killings (robust variables only)}
\label{tab:eba1}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk.pdf}
    \caption{Extreme Bounds Analysis -- Mass killings}
    \label{fig:hist-mk}
\end{sidewaysfigure}
\clearpage

%%%%% Add models during civil wars

@@

The following table shows the results of an EBA model using 3 variable combinations at a time. The results are very similar to those reported in the paper, in which we used 4 variables in each model.

To speed up computational time, the models included in this appendix are based on 50,000 random draws from the full set of all possible regression models. \citet[819]{salaimartin2004determinants} argue that random sampling produces unbiased estimates of the regression coefficients, and as such can accelerate convergence of the results. The models presented in the paper, however, include the full set of regressions.

@@

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-3vars.pdf}
    \caption{EBA -- 3 variables in each regression model}
    \label{fig:hist-mk-3vars}
\end{sidewaysfigure}
\clearpage

Table \ref{tab:mk-5vars} presents the results for models with up 5 variables in each regressions. In contrast with the main EBA model, the indicators of UCDP ongoing civil wars, ethnic diversity, and Polity IV square drop out of significance. Their CDF(0) is 0.88, just marginally below our threshold of 0.9.

@@

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-5vars.pdf}
    \caption{EBA -- 5 variables in each regression model}
    \label{fig:hist-mk-5vars}
\end{sidewaysfigure}
\clearpage

@@

In this section, we estimate the EBA model with different values of Variance Inflation Factor (VIF), a measure of multicollinearity for linear models. There is no standard definition about what constitutes an acceptable VIF value, although researchers often use a value of 10 as rule of thumb to indicate strong multicollinearity \citep[674]{o2007caution}. Our original model used 7 as a cutoff, and here we employ 10 (less strict), a lower value of 2.5 (more conservative), and a model without any VIF value (no restriction). The results are essentially identical to those in the main model. In the model with no VIF restriction, however, ethnic fractionalisation fails to meet our threshold by a very small margin. The CDF(0) of that covariate is 0.897, very close to the required value of 0.9.

@@

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{Base variables} &  &  &  &  &  \\
Log GDP per capita & -0.0091 & 0.0052 & 76.354 & 0.9343 & 50000 \\
 &  &  &  &  &  \\
\textit{Additional variables} &  &  &  &  &  \\
Post-Cold War years & -0.0134 & 0.0084 & 73.540 & 0.9495 & 7929 \\
UCDP civil war onset & 0.0529 & 0.0322 & 52.141 & 0.9438 & 4553 \\
Previous riots & 0.0140 & 0.0100 & 56.433 & 0.9216 & 7772 \\
UCDP ongoing civil war & 0.0172 & 0.0113 & 66.013 & 0.9113 & 4587 \\
Ethnic diversity (ELF) & 0.0182 & 0.0136 & 56.872 & 0.9056 & 8076 \\
Polity IV squared & -0.0002 & 0.0001 & 60.791 & 0.9021 & 7835 \\ \hline
\end{tabular}
\caption{EBA -- VIF 10 (less strict)}
\label{tab:eba1}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-high-vif.pdf}
    \caption{EBA -- VIF 10 (less strict)}
    \label{fig:hist-mk-high-vif}
\end{sidewaysfigure}
\clearpage

@@

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{Base variables} &  &  &  &  &  \\
Log GDP per capita & -0.0090 & 0.0051 & 76.055 & 0.9343 & 49620 \\
 &  &  &  &  &  \\
\textit{Additional variables} &  &  &  &  &  \\
Post-Cold War years & -0.0132 & 0.0084 & 72.845 & 0.9490 & 7929 \\
UCDP civil war onset & 0.0529 & 0.0322 & 52.378 & 0.9438 & 4553 \\
Previous riots & 0.0141 & 0.0101 & 56.242 & 0.9199 & 7772 \\
UCDP ongoing civil war & 0.0174 & 0.0114 & 65.652 & 0.9103 & 4587 \\
Ethnic diversity (ELF) & 0.0184 & 0.0137 & 56.674 & 0.9054 & 8076 \\
Polity IV squared & -0.0002 & 0.0001 & 61.206 & 0.90267 & 7835 \\ \hline
\end{tabular}
\caption{EBA -- VIF 2.5 (more conservative)}
\label{tab:eba1}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-high-vif.pdf}
    \caption{EBA -- VIF 2.5 (more conservative)}
    \label{fig:hist-mk-high-vif}
\end{sidewaysfigure}
\clearpage

>>>>

Our analysis tests the sensitivity of 40 variables on a sample of 177 countries from 1945 to 2013. We find that GDP \textit{per capita}, stable political regimes, and the post-Cold War period are negatively associated with mass killings. Conversely, previous political turmoil, ethnic diversity, and civil wars increase the risk of such violence. The distributed random forest algorithm indicates that years since the last episode of mass violence, GDP \textit{per capita}, urban population, ethnic polarisation, the number of military personnel, and democracy make the greatest contribution to our models' out-of-sample predictive power. While we discuss the importance of these findings in more detail later in this article, a few notes are imperative to make here. First, in some ways, these findings confirm previous research–unstable countries are more likely to witness the regime employ atrocity \citep[e.g.,]{goldsmith2013forecasting,harff2003no,krain1997state}.  However, many of the factors scholars often cite as observable indicators of such instability--regime transitions, coups d'état, the presence of militias, etc.--are not good proxies for instability. Thus, policymakers may be looking for the incorrect signs of impending atrocity when seeking to prevent its onset. Second, these findings raise concerns about policy options for preventing violence against civilians. If our conclusion is that unstable countries are violent, then preventing atrocity likely requires significant investments of time and resources in state-building, which is often politically and practically unfeasible \citep{doyle2006making}. Lastly, our findings are robust to a myriad of specifications, functional forms, and additional tests as we describe below. This analysis contributes significantly to the political violence literature by highlighting the parsimonious nature of the logic behind government atrocity and clearing away much of the empirical clutter surrounding this conclusion.

@@

We employ two methods to test the robustness of the potential determinants of state-led violence. First, we use a variant of extreme bounds analysis (EBA) to check the statistical significance of 35 explanatory variables cited in the academic literature.\footnote{The complete list of variables included in this article is available in appendix \ref{sec:mk-appendix}. Our selection procedure was as follows. We included all variables that appeared in at least two two quantitative papers on mass killings which employed logit or probit models in a global sample. We considered only published articles in our list.} Researchers have employed the EBA to assess the sensitivity of the determinants of civil war \citep{hegre2006sensitivity}, coups d'état \citep{gassebner2016expect}, democratisation \citep{gassebner2013extreme}, economic growth \citep{levine1992sensitivity, sala1997just, sturm2005determinants}, nuclear deterrence \citep{bell2015examining}, political repression \citep{hafner2005right}, and state corruption \citep{serra2006empirical}. The method is particularly useful when there is no consensus about which covariates belong in the ``true'' regression model \citep[178]{sala1997just} and scholars worry that omitted or unnecessary predictors could bias the parameter estimates \citep[60]{angrist2008mostly, clarke2005phantom, elwert2014endogenous, spector2011methodological}. By combining a large set of regression models into a single posterior distribution, the EBA reduces the uncertainty associated with the \textit{ad hoc} selection of predictors and provides an intuitive way of assessing the relative strength of the statistical findings \citep{leamer1985sensitivity, sala1997just}.

@@

In this paper, we follow his advice and consider the whole range of values of $CDF(0)$. We choose to use the whole distribution because the aggregate $CDF(0)$ allows us to move away from a binary indicator of robustness, the one we obtain with Leamer's extreme bounds, and present the estimations with their appropriate degrees of confidence \citep[179]{sala1997just}\footnote{As \citet[179]{sala1997just} describes, ``if 95 percent of the density function for the estimates of $\beta_{1}$ lies to the right of zero and only 52 percent of the density function for $\beta_{2}$  lies to the right of zero, one will probably think of variable 1 as being more likely to be correlated with [a dependent variable $Y$] than variable 2.''}. Our main focus is the percentage of the variable's cumulative distribution function that is smaller or greater than zero. We do not assume that the CDF has a normal distribution: Many of our coefficients do not follow a standard distribution, so the generic model provides a better fit to our data.\footnote{Histograms for all coefficients are available in appendix \ref{sec:mk-appendix}.} We specify our models as follows:

@@

A different set of variables reach significance if we only consider mass killings that occur during civil wars. We ran the models on three indicators of violent armed conflict. The first is the civil war measure by the Uppsala Conflict Database Program  (\citeyear{allansson2017organized,gleditsch2002armed}). Here, two variables meet our criteria for robustness, Post-Cold War years and conflicts with territorial aims, both negatively correlated with mass killing onset. When we limit our analysis to cases marked as civil wars by the Correlates of War project \citep{sarkees2010resort}, years since the last mass killing episode, previous riots, former human rights abuses, and ethnic diversity appear have a positive effect on the outcome. Rather surprisingly, the presence of militias has a negative effect on genocides and politicides. Finally, our model using ethnic civil war \citep{cederman2010ethnic} shows similar results: Militias have a negative impact on the likelihood of mass killings during ethnic conflict, as do conflicts over territorial aims. This finding is in contradiction with previous research \citep{koren2017means} and is deserving of further exploration. Overall, these results suggest two takeaways. First, the determinants of mass killing onset during peacetime and wartime are different from each other. While there are more restraints on atrocities during peacetime, these restraints fail once fighting begins. This raises particularly interesting questions about the pacifying nature of democracy. Second, how civil wars are classified significantly alters the results produced. This raises concerns for scholars about making claims based on only one data set.

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{UCDP data} &  &  &  &  &  \\
Territory aims & -0.044 & 0.019 & 74.997 & 0.9804 & 17902 \\
Post-Cold War years & -0.038 & 0.019 & 66.574 & 0.9222 & 17902 \\
 &  &  &  &  &  \\
\textit{COW data} &  &  &  &  &  \\
Physical integrity & 0.024 & 0.013 & 66.674 & 0.9564 & 17902 \\
Militias & -0.099 & 0.048 & 73.104 & 0.9490 & 17902 \\
Years since last mass killing & 0.006 & 0.002 & 88.208 & 0.9472 & 101583 \\
Previous riots & 0.078 & 0.041 & 65.412 & 0.9348 & 17902 \\
Ethnic diversity (ELF) & 0.095 & 0.062 & 48.615 & 0.9000 & 17902 \\
 &  &  &  &  &  \\
\textit{Cederman et al. data} &  &  &  &  &  \\
Territory aims & -0.051 & 0.026 & 74.288 & 0.9167 & 17902 \\
Militias & -0.050 & 0.035 & 52.240 & 0.9101 & 17902 \\ \hline
\end{tabular}
\caption{EBA -- Mass killings during civil wars (robust variables only)}
\label{tab:ucdp}
\end{table}

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=\textwidth, height=9cm]{images/drfdpp4.pdf}
\caption{Partial Dependence Plots -- Mass Killings during Civil Wars (Cederman et al. data)}
\label{fig:drfdpp4}
\end{center}
\end{figure}

@@

As a last set of robustness models, we estimate the same regressions using Harff's (\citeyear{harff2003no}) indicator of genocide and politicide. No variable appear as significant in our EBA models for genocide or politicide onset in peacetime. When we limit our sample to civil war years, the Post-Cold War period is again negatively correlated with the outcome when using the Correlates of War data set. Additionally, excluded population has a negative sign in more than 90\% of the models using both Correlates of War’s and Cederman et al's (\citeyear{cederman2010ethnic}) indicators of conflict. Displaced population also has a negative effect in the Correlates of War data set. During ethnic conflicts, our dummy variable for political assassinations has a negative impact on the onset of genocides. Overall, we conclude then that the significant covariate of genocide and politicide onset differ significantly from those of more general forms of government mass violence. Finally, the machine learning models are comparable to the ones we present above, with a similar set of variables appearing in the random forest estimations. The results are available in appendix \ref{sec:mk-appendix}.

@@

\subsubsection{Main Model}

Here we present a series of histograms with the coefficients' distribution of all variables in the main EBA model. There are 36 variables in total, seven of which are robust: Log GDP per capita, post-Cold War period, onset and ongoing civil wars (measured by the UCDP), previous riots, ethnic diversity and the squared term of the Polity IV index.

@@

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{Base variables} &  &  &  &  &  \\
Log GDP per capita & -0.0091 & 0.0052 & 76.055 & 0.9335 & 226707 \\
 &  &  &  &  &  \\
\textit{Additional variables} &  &  &  &  &  \\
Post-Cold War years & -0.0133 & 0.0085 & 72.845 & 0.9472 & 35614 \\
UCDP civil war onset & 0.0529 & 0.0321 & 52.378 & 0.9441 & 20854 \\
Previous riots & 0.0140 & 0.0100 & 56.242 & 0.9216 & 35614 \\
UCDP ongoing civil war & 0.0172 & 0.0115 & 65.652 & 0.9092 & 20854 \\
Ethnic diversity (ELF) & 0.0184 & 0.0137 & 56.674 & 0.9050 & 35614 \\
Polity IV squared & -0.0002 & 0.0001 & 61.206 & 0.9031 & 35614 \\ \hline
\end{tabular}
\caption{Extreme Bounds Analysis -- Mass killings}
\label{tab:mk}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk.pdf}
    \caption{Extreme Bounds Analysis -- Mass killings}
    \label{fig:mk}
\end{sidewaysfigure}
\clearpage

\subsection{Genocides during Civil Wars}
\label{sec:civil-wars}

Next, we discuss genocides that occur during wartime. We use three covariates that denote ongoing civil conflicts: one by the Uppsala Conflict Data Program \citep{allansson2017organized,gleditsch2002armed}, another by the Correlates of War \citep{sarkees2010resort}, and a third indicating the onset of ethnic conflict as coded by \citet{cederman2010ethnic}. The significant variables in the models below is notably different from the previous ones. We see that as evidence that genocides that take place during conflicts follow a particular logic and are not caused by the same factors that foster mass killings in peacetime.

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{UCDP data} &  &  &  &  &  \\
Territory aims & -0.044 & 0.019 & 74.997 & 0.9804 & 17902 \\
Post-Cold War years & -0.038 & 0.019 & 66.574 & 0.9222 & 17902 \\
 &  &  &  &  &  \\
\textit{COW data} &  &  &  &  &  \\
Physical integrity & 0.024 & 0.013 & 66.674 & 0.9564 & 17902 \\
Militias & -0.099 & 0.048 & 73.104 & 0.9490 & 17902 \\
Years since last mass killing & 0.006 & 0.002 & 88.208 & 0.9472 & 101583 \\
Previous riots & 0.078 & 0.041 & 65.412 & 0.9348 & 17902 \\
Ethnic diversity (ELF) & 0.095 & 0.062 & 48.615 & 0.9000 & 17902 \\
 &  &  &  &  &  \\
\textit{Cederman et al. data} &  &  &  &  &  \\
Territory aims & -0.051 & 0.026 & 74.288 & 0.9167 & 17902 \\
Militias & -0.050 & 0.035 & 52.240 & 0.9101 & 17902 \\ \hline
\end{tabular}
\caption{EBA -- Mass killings during civil wars}
\label{tab:ucdp1}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-ucdp.pdf}
    \caption{EBA -- Mass killings during civil wars (UCDP data)}
    \label{fig:mk-ucdp}
\end{sidewaysfigure}
\clearpage

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-ucdp.pdf}
    \caption{EBA -- Mass killings during civil wars (COW data)}
    \label{fig:mk-cow}
\end{sidewaysfigure}
\clearpage

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-ucdp.pdf}
    \caption{EBA -- Mass killings during ethnic civil wars (Cederman et al. data)}
    \label{fig:mk-eth}
\end{sidewaysfigure}
\clearpage

@@

The following table shows the results of an EBA model using 3 variable combinations per model. The results are very similar to those reported in the paper, in which we used 4 variables in each model.

The models below are based on 50,000 random draws from the full set of all possible regression models. \citet[819]{salaimartin2004determinants} argue that random sampling produces unbiased estimates of the regression coefficients with low computational time. The models presented in the manuscript, however, include the full set of possible regressions.

@@

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-3vars.pdf}
    \caption{EBA -- 3 variables in each regression model}
    \label{fig:mk-3vars}
\end{sidewaysfigure}
\clearpage

Table \ref{tab:mk-5vars} presents the results for models with up 5 variables in each regressions. In contrast with the main EBA model, the indicators of UCDP ongoing civil wars, ethnic diversity, and Polity IV square drop out of significance. Their individual CDFs(0) are about 0.88, just marginally below our specified threshold of 0.9.

@@

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-5vars.pdf}
    \caption{EBA -- 5 variables in each regression model}
    \label{fig:mk-5vars}
\end{sidewaysfigure}
\clearpage

@@

In this section, we estimate the EBA model with different values of Variance Inflation Factor (VIF), a measure of multicollinearity for linear models. There is no standard definition about what constitutes an acceptable VIF value, although researchers often use a value of 10 as rule of thumb to indicate strong multicollinearity \citep[674]{o2007caution}. Our original model used a slightly more conservative value of 7 as a cutoff. Here we test the same model with VIF $=$ 10 (less strict), 2.5 (more conservative), and a model without any VIF value (no restriction). The results are essentially identical to those in the main model. In the model with no VIF restriction, however, ethnic fractionalisation fails to meet our threshold by a very small margin. The CDF(0) of that covariate is 0.897, very close to the required value of 0.9.

@@

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{Base variables} &  &  &  &  &  \\
Log GDP per capita & -0.0091 & 0.0052 & 76.354 & 0.9343 & 50000 \\
 &  &  &  &  &  \\
\textit{Additional variables} &  &  &  &  &  \\
Post-Cold War years & -0.0134 & 0.0084 & 73.540 & 0.9495 & 7929 \\
UCDP civil war onset & 0.0529 & 0.0322 & 52.141 & 0.9438 & 4553 \\
Previous riots & 0.0140 & 0.0100 & 56.433 & 0.9216 & 7772 \\
UCDP ongoing civil war & 0.0172 & 0.0113 & 66.013 & 0.9113 & 4587 \\
Ethnic diversity (ELF) & 0.0182 & 0.0136 & 56.872 & 0.9056 & 8076 \\
Polity IV squared & -0.0002 & 0.0001 & 60.791 & 0.9021 & 7835 \\ \hline
\end{tabular}
\caption{EBA -- VIF 10}
\label{tab:mk-high-vif}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-high-vif.pdf}
    \caption{EBA -- VIF 10}
    \label{fig:mk-high-vif}
\end{sidewaysfigure}
\clearpage

@@

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{Base variables} &  &  &  &  &  \\
Log GDP per capita & -0.0090 & 0.0051 & 76.055 & 0.9343 & 49620 \\
 &  &  &  &  &  \\
\textit{Additional variables} &  &  &  &  &  \\
Post-Cold War years & -0.0132 & 0.0084 & 72.845 & 0.9490 & 7929 \\
UCDP civil war onset & 0.0529 & 0.0322 & 52.378 & 0.9438 & 4553 \\
Previous riots & 0.0141 & 0.0101 & 56.242 & 0.9199 & 7772 \\
UCDP ongoing civil war & 0.0174 & 0.0114 & 65.652 & 0.9103 & 4587 \\
Ethnic diversity (ELF) & 0.0184 & 0.0137 & 56.674 & 0.9054 & 8076 \\
Polity IV squared & -0.0002 & 0.0001 & 61.206 & 0.90267 & 7835 \\ \hline
\end{tabular}
\caption{EBA -- VIF 2.5}
\label{tab:low-vif}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-high-vif.pdf}
    \caption{EBA -- VIF 2.5}
    \label{fig:mk-low-vif}
\end{sidewaysfigure}
\clearpage

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{Base variables} &  &  &  &  &  \\
Log GDP per capita & -0.0091 & 0.0052 & 75.940 & 0.9343 & 50000 \\
 &  &  &  &  &  \\
\textit{Additional variables} &  &  &  &  &  \\
Post-Cold War years & -0.0133 & 0.0085 & 72.756 & 0.9469 & 7800 \\
UCDP civil war onset & 0.0531 & 0.0321 & 53.068 & 0.9452 & 4596 \\
Previous riots & 0.0140 & 0.0101 & 56.139 & 0.9200 & 7811 \\
UCDP ongoing civil war & 0.0170 & 0.0116 & 64.487 & 0.9057 & 4497 \\
Ethnic diversity (ELF) & 0.0184 & 0.0137 & 56.814 & 0.9056 & 7808 \\
Polity IV squared & -0.0002 & 0.0001 & 60.825 & 0.9009 & 7903 \\ \hline
\end{tabular}
\caption{EBA -- No VIF Restriction}
\label{tab:mk-no-vif}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-no-vif.pdf}
    \caption{EBA -- No VIF restriction}
    \label{fig:mk-no-vif}
\end{sidewaysfigure}
\clearpage

We estimate the same EBA with logit and probit in addition to the linear probability models reported above. Logistic and probit regressions may have issues of complete separation, that is, when some covariates perfectly separates zeros and ones in the outcome variable.\footnote{We thank Mark Bell for sharing \texttt{R} code to estimate penalised-likelihood models.} In that case, the estimations fail to converge. We addressed this problem by adding a weak prior to the regression coefficients as suggested by \citet{gelman2008weakly}. We first scaled the non-binary variables to have a mean of 0 and a standard deviation of 0.5, then we added a Cauchy distribution with centre 0 and scale 2.5. The probit regressions use a scale of $2.5 \times 1.6$, which is also recommended by the authors \citep{arm2017rpackage}. Ethnic diversity and ongoing civil wars come close to meeting our threshold values (0.88 and 0.84, respectively), and civil war onset (UCDP) has a higher percentage of significant coefficients and a high CDF(0) area than in the linear probability models.

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{Base variables} &  &  &  &  &  \\
Log GDP per capita & 0.434 & 0.223 & 75.570 & 0.9267 & 50000 \\
 &  &  &  &  &  \\
\textit{Additional variables} &  &  &  &  &  \\
UCDP civil war onset & 1.308 & 0.530 & 87.261 & 0.9742 & 4506 \\
Post-Cold War years & -0.911 & 0.428 & 70.456 & 0.9448 & 7890 \\
Previous riots & 0.744 & 0.38 & 66.778 & 0.9383 & 7805 \\
Polity IV squared & -0.015 & 0.008 & 68.038 & 0.9285 & 7975 \\ \hline
\end{tabular}
\caption{EBA -- Logistic model}
\label{tab:mk-logit}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-logit.pdf}
    \caption{EBA -- Logistic model}
    \label{fig:mk-logit}
\end{sidewaysfigure}
\clearpage

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\hline
\textbf{Variable} & \textbf{Avg. $\beta$} & \textbf{Avg. SE} & \textbf{$\%$ Sig.} & \textbf{CDF(0)} & \textbf{Models} \\ \hline
\textit{Base variables} &  &  &  &  &  \\
Log GDP per capita & -0.1924 & 0.1031 & 76.118 & 0.9258 & 50000 \\
 &  &  &  &  &  \\
\textit{Additional variables} &  &  &  &  &  \\
UCDP civil war onset & 0.6422 & 0.2582 & 89.225 & 0.9772 & 4501 \\
Previous riots & 0.3367 & 0.1743 & 71.813 & 0.9436 & 7851 \\
Post-Cold War years & -0.3709 & 0.1830 & 71.465 & 0.9404 & 7836 \\
Polity IV squared & -0.0061 & 0.0032 & 70.155 & 0.9315 & 7931 \\ \hline
\end{tabular}
\caption{EBA -- Probit model}
\label{tab:eba1}
\end{table}

\clearpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=\textwidth]{images/mk-probit.pdf}
    \caption{EBA -- Probit model}
    \label{fig:mk-probit}
\end{sidewaysfigure}
\clearpage

In the manuscript, we perform a grid search to optimise the hyperparameters of the random forest models. Because the grid search evaluates a wide range of parameter values at once, it is often unnecessary to run additional tests to assess the robustness of the results. Nevertheless, as random forests themselves are an approximation to a myriad of possible parameter combinations, changes in seed numbers may influence model output. Thus, we start the models with different random seed numbers to evaluate how study are our original results.\footnote{The numbers were generated at \href{https://www.random.org/}{https://www.random.org/}.} The findings holds quite well: Although variable importance chances from one model to another, the most significant variables appear repeatedly in the estimations. The marginal plots also show that their effects on the outcome variable are similar despite the nonlinearities. Below we show the 6 most significant variables and their respective partial dependence plots.
